---
title: "Dodgers"
author: "John Zelten"
date: "2022-10-26"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = "", warning = FALSE, message = FALSE)
```

# Los Angeles Dodgers 2022 Pitching Graphs

In additon to my project I wanted to incorporate some graphs that could be used for our own players performance rather than scouting other teams players. I can use these codes to do the same for opposition pitching, would just need to change the filters. 

```{r}


library(ggplot2) # Load ggplot2
library(ggdark) # Load ggdark
library(baseballr) # Load baseballr package
library(tidyr) # For gather function
library(ggnewscale) # For multiple color scales


```

```{r}
load("~/Portfolio Projects/2022_stat_cast_data.rda")
```


#Pitch type by count
Using these graphs we can see our pitchers tendencies on what kind of pitches are thrown in certain counts. 

```{r}
pitchers <- pitchers

count_diff <- rep("equal", nrow(pitchers))
count_diff[pitchers$balls < pitchers$strikes] <- "ahead"
count_diff[pitchers$balls > pitchers$strikes] <- "behind"
pitchers$count_diff <- count_diff

# Extract rows with pitchers from Selected Team
pitchers_1 <- pitchers[pitchers$pitcher_team == "LAD",]
players_1 <- unique(na.omit(pitchers_1$player_name))
pitches <- rep(0, length(players_1))
for(i in 1:length(players_1)){
    pitches[i] <- sum(pitchers_1$player_name == players_1[i], na.rm = T)
}
temp_1 <- cbind.data.frame(players_1[order(pitches, decreasing = TRUE)],
                            pitches[order(pitches, decreasing = TRUE)])
names(temp_1) <- c("players", "pitches")

sel_players <- c(1:5)

for(i in sel_players){
  player <- temp_1$players[i]
  t1 <- as.matrix(table(pitchers$pitch_name[pitchers$player_name == player], pitchers$count_diff[pitchers$player_name == player]))
  p_sums <- rowSums(t1)
  rownames(t1) <- paste(rownames(t1), ": ", p_sums, sep = "")
  temp <- colSums(t1)
  t1 <- as.data.frame(t1)  
  t1$Freq[t1$Var2 == "ahead"] <- t1$Freq[t1$Var2 == "ahead"]/temp[1]
  t1$Freq[t1$Var2 == "behind"] <- t1$Freq[t1$Var2 == "behind"]/temp[2]
  t1$Freq[t1$Var2 == "equal"] <- t1$Freq[t1$Var2 == "equal"]/temp[3]
  names(t1)[1:2] <- c("pitch", "count")
  g_1 <- ggplot(t1, 
              aes(x = count, y = pitch)) +
  geom_tile(aes(fill = Freq)) + 
  scale_fill_gradient(low = "blue",  
                       high = "red") +
  labs(x = "Count Status", 
       y = "Pitch",
       fill = "Proportion by\nCount Status",
       title = paste(player, "- Pitches:", temp_1$pitches[i], sep = ""),
       subtitle = "Pitch Selection\nby Count")

 print(g_1)
  
}
```

# Ball movement

These graphs will show the flight path of the ball from the pitchers hand to when it crosses the plate.

```{r}

sel_players <- c(1:5)

for(i in sel_players){ 
   player <- temp_1$players[i]
    player_data <- pitchers[pitchers$player_name == player,]
g_1 <- ggplot(player_data[player_data$description %in% c("called_strike", "hit_into_play",
                                                         "hit_into_play_no_out", "hit_into_play_score",
                                                         "swinging_strike"),], 
              aes(x = release_pos_x, 
                  xend = plate_x, 
                  y = release_pos_z,
                  yend = plate_z, 
                  color = pitch_name)) + 

  geom_segment(alpha = 0.3,
               arrow = arrow(length = unit(0.25, "cm"), 
                                         type = "closed")) +
  dark_theme_bw() + 
  geom_vline(xintercept = -0.7083333, linetype = 2) + 
  geom_vline(xintercept = 0.7083333, linetype = 2) + 
  #scale_color_gradient(low = "blue", high= "red") +
  geom_point(inherit.aes = FALSE, aes(x = release_pos_x, 
                                      y= release_pos_z, 
                                      color = pitch_name), 
             alpha = 0.5)  +
  labs(x = "X", y = "Z",  
       title = paste(player),
       subtitle = "Pitch Movement by Pitch Type 2022")
print(g_1)

}


```



